---
title: "LSGF_markdown"
author: "Angela Jiang"
date: '2022-10-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
```

## R Markdown

```{r read_file}
table = read.table(file = 'Orthogroups.GeneCount-new_taxa-10-23.tsv', sep = '\t', header = TRUE)

taxon = read.table(file = 'taxon.csv', sep = ',', header = TRUE)
```


```{r table, echo=FALSE}
helperFunction1 <- function(x){
  ifelse(x>=1, 1,0)
}

reduced_table <- table %>%
  filter(Total > 5)

reduced_table_with_row_name <- reduced_table[,-1]
rownames(reduced_table_with_row_name) <- reduced_table[,1]

newtable <- reduced_table %>%
  filter(Total > 5) %>%
  mutate(across(!Orthogroup, helperFunction1)) %>%
  select(!228) 

newtablenew <- newtable[,-1]
rownames(newtablenew) <- newtable[,1]

colnames(newtablenew) <- taxon$Group

OGs <- newtable %>%
  select(contains("Orthogroup"))

# sums columns with the same name and merges them
by_group <- as.data.frame(sapply(unique(names(newtablenew)), function(i)rowSums(newtablenew[names(newtablenew) == i])))
```

## Arcellinida-Specific genes

```{r arcellinida_specific}
arcellinidaonly <- by_group %>%
  filter(Group_1 > 1 & Group_2 > 2 & Group_3 > 2 & Group_4 > 1 & Group_5 > 0 & Group_6 > 0 & Group_8 > 3 & Group_7/(Group_1+Group_2+Group_3+Group_4+Group_5+Group_6+Group_8+Group_7) < 0.05)

#saving to table 
arcellinida_OGs <- merge(reduced_table_with_row_name, arcellinidaonly,
                          by = 'row.names', all = FALSE)[1:227]

names(arcellinida_OGs)[names(arcellinida_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(arcellinida_OGs, sep="\t", file="Tables/LSGarcellinida.tsv", row.names = FALSE)

#calculate number of LSG
number <- nrow(arcellinidaonly)

```

Criteria: at least 2 from Sphaerothecina, 3 H. elegans, 3 Nebela, 2 Difflugina, 1 Heleopera, 1 from misc Arcellinida groups, and less than 5% non Arcellinida

There are `r number` all-Arcellinida lineage-specific genes.

## H. papilio-Specific gene families

```{r pap_specific}
hpapilio_only <- by_group %>%
  filter(Group_8 > 20 & ((Group_8/(Group_1+Group_2+Group_3+Group_4+Group_5+Group_6+Group_7+Group_8)) > 0.95))

#saving to table 
hpapilio_OGs <- merge(reduced_table_with_row_name, hpapilio_only,
                          by = 'row.names', all = FALSE)[1:227]

#rename column name
names(hpapilio_OGs)[names(hpapilio_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(hpapilio_OGs, sep="\t", file="Tables/LSGpapilio.tsv", row.names = FALSE)

#calculate number of LSG
number_pap <- nrow(hpapilio_only)

```

Criteria: at least 20 H. papilio cells, 5% of other cells can be non-H. papilio

There are `r number_pap` H. papilio-only lineage-specific genes.

## H. elegans-Specific gene families

```{r elegans_specific}
helegans_only <- by_group %>%
  filter(Group_2 > 10 & ((Group_2/(Group_1+Group_2+Group_3+Group_4+Group_5+Group_6+Group_7+Group_8)) > 0.95))

#saving to table 
elegans_OGs <- merge(reduced_table_with_row_name, helegans_only,
                          by = 'row.names', all = FALSE)[1:227]

#rename column name
names(elegans_OGs)[names(elegans_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(elegans_OGs, sep="\t", file="Tables/LSGelegans.tsv", row.names = FALSE)

#calculate number of LSG
number_elegans <- nrow(helegans_only)

```

Criteria: at least 10 H. elegans cells, 5% of other cells can be non-H. elegans

There are `r number_elegans` H. elegans-only lineage-specific genes.

## Heatmaps!

### All Arcellinida Heatmap
```{r heatmap_arc}
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
arcellinida_heat <- merge(newtable_heat, arcellinidaonly,
                          by = 'row.names', all = FALSE)[1:227]

#rename column name
names(arcellinida_heat)[names(arcellinida_heat) == 'Row.names'] <- 'Orthogroup'
arcellinida_heat_x <- arcellinida_heat[,-1]
rownames(arcellinida_heat_x) <- arcellinida_heat[,1]

data <- as.matrix(arcellinida_heat_x)
fig <- plot_ly(z = data, y=rownames(arcellinida_heat_x), x=colnames(arcellinida_heat_x), colors = colorRamp(c("white", "red")), type = "heatmap") %>%
  layout( plot_bgcolor='#e5ecf6',
          yaxis = list(
            showticklabels = FALSE,
            gridcolor = 'ffff'))#generate an interactive heat map
fig #view

```

### H. papilio LSG Heatmap
```{r heatmap_pap}
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
papilio_heat <- merge(newtable_heat, hpapilio_only,
                          by = 'row.names', all = FALSE)[1:227]

#rename column name
names(papilio_heat)[names(papilio_heat) == 'Row.names'] <- 'Orthogroup'
papilio_heat_x <- papilio_heat[,-1]
rownames(papilio_heat_x) <- papilio_heat[,1]

data <- as.matrix(papilio_heat_x)
fig <- plot_ly(z = data, y=rownames(papilio_heat_x), x=colnames(papilio_heat_x), colors = colorRamp(c("white", "red")), type = "heatmap") %>%
  layout( plot_bgcolor='#e5ecf6',
          yaxis = list(
            showticklabels = FALSE,
            gridcolor = 'ffff'))#generate an interactive heat map
fig #view

```

### H. elegans LSG Heatmap
```{r heatmap_elegans}
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
elegans_heat <- merge(newtable_heat, helegans_only,
                          by = 'row.names', all = FALSE)[1:227]

#rename column name
names(elegans_heat)[names(elegans_heat) == 'Row.names'] <- 'Orthogroup'
elegans_heat_x <- elegans_heat[,-1]
rownames(elegans_heat_x) <- elegans_heat[,1]

data <- as.matrix(elegans_heat_x)
fig <- plot_ly(z = data, y=rownames(elegans_heat_x), x=colnames(elegans_heat_x), colors = colorRamp(c("white", "red")), type = "heatmap") %>%
  layout( plot_bgcolor='#e5ecf6',
          yaxis = list(
            showticklabels = FALSE,
            gridcolor = 'ffff'))#generate an interactive heat map
fig #view

```
