---
title: "LSGF_markdown"
author: "Angela Jiang"
date: '2023-4-11'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(plotly)
library(processx)
library(ggplot2)
```

Given an OrthoFinder gene count table, finds Arcellinida, H. elegans and H. papilio LSGFs

## Reads the Tables

```{r read_file}
table = read.table(file = 'curated_genecount_elegans.csv', sep = ',', header = TRUE)

taxon = read.table(file = 'taxon-tides.csv', sep = ',', header = TRUE)
```

## Sets Up the Dataframes

```{r table, echo=FALSE}
helperFunction1 <- function(x){
  ifelse(x>=1, 1,0)
}

reduced_table <- table
 #filter(Total > 5)

reduced_table_with_row_name <- reduced_table[,-1]
rownames(reduced_table_with_row_name) <- reduced_table[,1]

newtable <- reduced_table %>%
  #filter(Total > 5) %>%
  subset(select = -c(Am_tu_He44) ) %>% #longest: Am_tu_He44.FinalORFs.AA
  #subset(select = -c(Am_tu_Hp19.FinalORFs.AA) ) %>%
  mutate(across(!Orthogroup, helperFunction1)) %>%
  select(!223) #224 for tides, 228 longest, 223 for curated data

newtablenew <- newtable[,-1]
rownames(newtablenew) <- newtable[,1]

colnames(newtablenew) <- taxon$Group

OGs <- newtable %>%
  select(contains("Orthogroup"))

# sums columns with the same name and merges them
by_group <- as.data.frame(sapply(unique(names(newtablenew)), function(i)rowSums(newtablenew[names(newtablenew) == i])))
```

## Arcellinida-Specific genes

```{r arcellinida_specific}
arcellinidaonly <- by_group %>%
  filter(Sphaerothecina > 1 & H_elegans > 2 & Nebela > 2 & Centropyxis_Difflugia_Lesquereusia > 1 & Misc_Arcellinida > 0 & H_papilio > 3 & Non_Arcellinida/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+H_papilio+Non_Arcellinida) < 0.05)
# ^ with taxon picky criteria

#arcellinidaonly <- by_group %>%
#  filter((Sphaerothecina + H_elegans + Nebela + Centropyxis_Difflugia_Lesquereusia + Misc_Arcellinida + H_papilio >= 12) & Non_Arcellinida/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+H_papilio+Non_Arcellinida) < 0.05)

#saving to table 
arcellinida_OGs <- merge(reduced_table_with_row_name, arcellinidaonly,
                          by = 'row.names', all = FALSE)[1:225]

names(arcellinida_OGs)[names(arcellinida_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(arcellinida_OGs, sep="\t", file="Tables/LSGarcellinida.tsv", row.names = FALSE)

#calculate number of LSG
number <- nrow(arcellinidaonly)

```

Criteria: at least 2 from Sphaerothecina, 3 H. elegans, 3 Nebela, 2 Difflugina, 1 Heleopera, 1 from misc Arcellinida groups, and less than 5% non Arcellinida

There are `r number` all-Arcellinida lineage-specific genes.

## H. papilio-Specific gene families

```{r pap_specific}
hpapilio_only <- by_group %>%
  filter(H_papilio > 20 & ((H_papilio/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+Non_Arcellinida+H_papilio)) > 0.95))

#saving to table 
hpapilio_OGs <- merge(reduced_table_with_row_name, hpapilio_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(hpapilio_OGs)[names(hpapilio_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(hpapilio_OGs, sep="\t", file="Tables/LSGpapilio.tsv", row.names = FALSE)

#calculate number of LSG
number_pap <- nrow(hpapilio_only)

```

Criteria: at least 20 H. papilio cells, 5% of other cells can be non-H. papilio

There are `r number_pap` H. papilio-only lineage-specific genes.

## Hyalospheniidae-Specific gene families

```{r hyalospheniidae}
hyalospheniidae_only <- by_group %>%
  filter(H_papilio > 3 & H_elegans > 3 & (((H_papilio + H_elegans)/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+Non_Arcellinida+H_papilio)) > 0.95))

#saving to table 
hyalospheniidae_OGs <- merge(reduced_table_with_row_name, hyalospheniidae_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(hyalospheniidae_OGs)[names(hyalospheniidae_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(hyalospheniidae_OGs, sep="\t", file="Tables/LSGhyalospheniidae.tsv", row.names = FALSE)

#calculate number of LSG
number_hyalospheniidae <- nrow(hyalospheniidae_only)

```

Criteria: at least 3 H. papilio cells, 3 H. elegans, 3 Nebela, 1 quadruella, 5% of other cells can be non-Hyalospheniidae

There are `r number_hyalospheniidae` Hyalospheniidae-only lineage-specific genes.

## Sphaerothecina-Specific gene families

```{r Sphaerothecina}
Sphaerothecina_only <- by_group %>%
  filter(Sphaerothecina > 5 & (((Sphaerothecina)/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+Non_Arcellinida+H_papilio)) > 0.95))

#saving to table 
Sphaerothecina_OGs <- merge(reduced_table_with_row_name, Sphaerothecina_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(Sphaerothecina_OGs)[names(Sphaerothecina_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(Sphaerothecina_OGs, sep="\t", file="Tables/LSG_Sphaerothecina.tsv", row.names = FALSE)

#calculate number of LSG
number_Sphaerothecina <- nrow(Sphaerothecina_only)

```

Criteria: at least 5 Sphaerothecina, 5% of other cells can be non-Sphaerothecina

There are `r number_Sphaerothecina` Sphaerothecina-only lineage-specific genes.

## Longithecina-Specific gene families

```{r Longithecina}
Longithecina_only <- by_group %>%
  filter((Sphaerothecina > 2 & Centropyxis_Difflugia_Lesquereusia > 2) & (((Sphaerothecina)/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+Non_Arcellinida+H_papilio)) > 0.95))

#saving to table 
Longithecina_OGs <- merge(reduced_table_with_row_name, Longithecina_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(Longithecina_OGs)[names(Longithecina_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(Longithecina_OGs, sep="\t", file="Tables/LSG_Longithecina.tsv", row.names = FALSE)

#calculate number of LSG
number_Longithecina <- nrow(Longithecina_only)

```

Criteria: at least 5 Longithecina, 5% of other cells can be non-Longithecina

There are `r number_Longithecina` Longithecina-only lineage-specific genes.

Note: None found that weren't found in Sphaerothecina-specific genes already

## H. elegans-Specific gene families

```{r elegans_specific}
helegans_only <- by_group %>%
  filter(H_elegans > 7 & ((H_elegans/(Sphaerothecina+H_elegans+Nebela+Centropyxis_Difflugia_Lesquereusia+Heleopera+Misc_Arcellinida+Non_Arcellinida+H_papilio)) > 0.95))

#saving to table 
elegans_OGs <- merge(reduced_table_with_row_name, helegans_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(elegans_OGs)[names(elegans_OGs) == 'Row.names'] <- 'Orthogroup'

write.table(elegans_OGs, sep="\t", file="Tables/LSGelegans.tsv", row.names = FALSE)

#calculate number of LSG
number_elegans <- nrow(helegans_only)

```

Criteria: at least 7 H. elegans cells, 5% of other cells can be non-H. elegans

There are `r number_elegans` H. elegans-only lineage-specific genes.

## Heatmaps!

### All Arcellinida Heatmap
```{r heatmap_arc, fig.width = 10, fig.height = 15}
library(reshape2)
library(pheatmap)
library(RColorBrewer)
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
arcellinida_heat <- merge(newtable_heat, arcellinidaonly,
                          by = 'row.names', all = FALSE)[1:224]

#rename column name
names(arcellinida_heat)[names(arcellinida_heat) == 'Row.names'] <- 'Orthogroup'
arcellinida_heat_x <- arcellinida_heat[,-1]
rownames(arcellinida_heat_x) <- arcellinida_heat[,1]

arcellinida_heat_x <- arcellinida_heat_x %>%
  select(contains("Am_tu_"))

data <- as.matrix(arcellinida_heat_x)
datasorted<-data[, order(colSums(-data))]

datasorted <- datasorted[order(rowSums(datasorted),decreasing=F),]


ggplotthing <- melt(data, na.rm = FALSE, value.name = "name", id = "columns")

p <- ggplot(ggplotthing, aes(Var1, Var2, fill= name)) + 
  geom_tile()
#try pheatmap()


taxonnew <- taxon
rownames(taxonnew) <- taxonnew[,1]
taxonnew <- taxonnew[c(2)]

row.names(taxonnew) <- str_trunc(row.names(taxonnew), 13, "right") 

colnames(datasorted) <- str_trunc(colnames(datasorted), 13, "right") 
clustered <- pheatmap(datasorted, color=colorRampPalette(c("white", "black"))(50), annotation_col = taxonnew, show_rownames = F, show_colnames = F)

#fig <- plot_ly(z = data, y=rownames(arcellinida_heat_x), #x=colnames(arcellinida_heat_x), colors = colorRamp(c("white", "red")), #type = "heatmap") %>%
#  layout( plot_bgcolor='#e5ecf6',
#          yaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff'),
#          xaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff')
#          )#generate an interactive heat map
#fig #view

```

### Dendrogram
```{r, fig.width = 30, fig.height = 30}
library(WGCNA)
library(ggdendro)
library(dendextend)
# Compute Euclidean distance between samples
dist=dist(t(datasorted), diag=TRUE)

# Perfor clustering with hclust
hc <- hclust(dist)

dhc <- as.dendrogram(hc)

colors_to_use <- taxonnew$Color
colors_to_use <- colors_to_use[order.dendrogram(dhc)]

labels_colors(dhc) <- colors_to_use
# Now each level of complexity has a color
#labels_colors(dend) #---> This will list the row numbers that appear on the dendrogram from left to right.  


#now we need to change the labels to the level names of complexity
dhc.list<-as.character(taxonnew$Group)

#now we need to change the order of labels from the order in which they appear in the original data to match the order in which they appear on the dendrogram we just made.
labels(dhc)<- dhc.list[order.dendrogram(dhc)]

dhc <- set(dhc, "labels_cex", 4)

#Here is the final plot!
plot(dhc, main = "Taxon", cex=1.5)

# Rectangular lines
#ddata <- dendro_data(dhc, type = "rectangle")
#p <- ggplot(segment(ddata)) + 
#  geom_segment(aes(x = x, y = y, xend = xend, yend = yend)) +
#  geom_text(data = label(ddata), 
#            aes(x = x, y = y, label = label, hjust = 0), 
#            size = 5) +
#  coord_flip() + scale_y_reverse(expand = c(0.2, 0)) + theme_classic()
#
#p

#dhc <- as.dendrogram(hc)
#dhc <- ggdendrogram(hc, rotate = FALSE, size = 4) + theme(text = element_text(size = 20))   

#plot(dhc, cex = 2)

```

### H. papilio LSG Heatmap
```{r heatmap_pap, fig.width = 10, fig.height = 10}
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
papilio_heat <- merge(newtable_heat, hpapilio_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(papilio_heat)[names(papilio_heat) == 'Row.names'] <- 'Orthogroup'
papilio_heat_x <- papilio_heat[,-1]
rownames(papilio_heat_x) <- papilio_heat[,1]

papilio_heat_x <- papilio_heat_x %>%
  select(contains("Am_tu_Hp"))


papilio_heat_x <- papilio_heat_x %>%
  subset(select = -c(1) )

data <- as.matrix(papilio_heat_x)

papilio_taxon_filtered <- subset(taxonnew, grepl("Am_tu_Hp", rownames(taxonnew)))
papilio_taxon_filtered <- subset(papilio_taxon_filtered, !(rownames(papilio_taxon_filtered) == "Am_tu_Hp.TIdeS_Pred.AA"))
row_names <- rownames(papilio_taxon_filtered)
df_ordered_papilio <- as.data.frame(papilio_taxon_filtered[order(papilio_taxon_filtered$Location), ])
rownames(df_ordered_papilio) <- rownames(papilio_taxon_filtered)[order(papilio_taxon_filtered$Location)]
colnames(df_ordered_papilio) <- c("Location")
#colnames(data) <- str_trunc(colnames(data), 13, "right") 


col_indices <- match(rownames(df_ordered_papilio), colnames(data))
data_ordered <- data[, col_indices]
data_ordered <- data_ordered[,-47]

#colnames(data) <- str_trunc(colnames(data), 13, "right")
clustered <- pheatmap(papilio_heat_x, color=colorRampPalette(c("white", "black"))(50), show_rownames = F)

#fig <- plot_ly(z = data, y=rownames(papilio_heat_x), #x=colnames(papilio_heat_x), colors = colorRamp(c("white", "red")), type #= "heatmap") %>%
#  layout( plot_bgcolor='#e5ecf6',
#          yaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff'),
#          xaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff'))#generate an interactive heat map
#fig #view

```

### H. elegans LSG Heatmap
```{r heatmap_elegans, fig.width = 10, fig.height = 10}
library(reshape)
newtable_heat <- newtable[,-1]
rownames(newtable_heat) <- newtable[,1]

#saving to binary table 
elegans_heat <- merge(newtable_heat, helegans_only,
                          by = 'row.names', all = FALSE)[1:225]

#rename column name
names(elegans_heat)[names(elegans_heat) == 'Row.names'] <- 'Orthogroup'
elegans_heat_x <- elegans_heat[,-1]
rownames(elegans_heat_x) <- elegans_heat[,1]

elegans_heat_x <- elegans_heat_x %>%
  select(contains("Am_tu_He"))
  #select(-c(Am_tu_He43))

data <- as.matrix(elegans_heat_x)
elegans_taxon_filtered <- subset(taxonnew, grepl("Am_tu_He", rownames(taxonnew)))
elegans_taxon_filtered <- subset(elegans_taxon_filtered, !(rownames(elegans_taxon_filtered) == "Am_tu_He44"))
row_names <- rownames(elegans_taxon_filtered)
df_ordered <- as.data.frame(elegans_taxon_filtered[order(elegans_taxon_filtered$Location), ])
rownames(df_ordered) <- rownames(elegans_taxon_filtered)[order(elegans_taxon_filtered$Location)]
colnames(df_ordered) <- c("Location")
#colnames(data) <- str_trunc(colnames(data), 13, "right") 


col_indices <- match(rownames(df_ordered), colnames(data))
data_ordered <- data[, col_indices]

clustered <- pheatmap(elegans_heat_x, color=colorRampPalette(c("white", "black"))(50), show_rownames = F, cluster_cols = T)


#fig <- plot_ly(z = data, y=rownames(elegans_heat_x), #x=colnames(elegans_heat_x), colors = colorRamp(c("white", "red")), type #= "heatmap") %>%
#  layout( plot_bgcolor='#e5ecf6',
#          yaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff'),
#          xaxis = list(
#            showticklabels = FALSE,
#            gridcolor = 'ffff'))#generate an interactive heat map
#fig #view

```
